---
title: "Exploratory Analysis of Trump's Speeches"
author: "Jan Schubert"
date: "12 1 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10)
library(data.table)
library(quanteda)
library(dendextend)
library(ggraph)
```

### About data
This data contains 863 transcripts of speeches of Donald Trump. This dataset was created by scraping Trump’s speeches from YouTube and using Google voice assistant. [Here is the data](https://www.kaggle.com/binksbiz/mrtrump)


### Cleaning
Steaming and stopwords
```{r}
### Load data
trump <- fread("../data/MrTrumpSpeeches.csv", stringsAsFactors = FALSE, sep = "~")
# trump <- trump[sample(1:nrow(trump), 100), ]
trump[, upload_date := as.Date(as.character(upload_date), format = "%Y%m%d")]
trump[, year := as.factor(format(upload_date, format = "%Y"))]

### ------------------------------
### Cleaning
trump_corpus <- corpus(trump$subtitles)
docvars(trump_corpus, "Year")<- trump$year
docvars(trump_corpus, "Title")<- trump$title

## tokenization and cleaning
toks <- 
  tokens(trump_corpus, remove_numbers = TRUE, remove_punct = TRUE, remove_symbols = TRUE, remove_separators = TRUE) %>%
  # tokens_wordstem %>%
  tokens_remove(c(stopwords(language = "en"), "going", "thanks"))
trump_dfm <- dfm(toks, verbose = TRUE)
```


### What are the most common words?
```{r}
# word frequency
textplot_wordcloud(trump_dfm, min_count  = 100)
barplot(topfeatures(trump_dfm), cex.names = .8, main = "10 most frequent words")
```


### Which words are used to talk about different topics?
Bigram to identify key topics in Trump's speeches. 
```{r}
## ------------------------------------------
## bigram
trump_dfm2 <- dfm(toks, ngrams = 2)
trump_dfm2 <- dfm_trim(trump_dfm2, min_termfreq = 10)
# textplot_wordcloud(trump_dfm2)
topfeatures(trump_dfm2)


trump_dtm2_dt <- as.data.table(convert(trump_dfm2, to = "data.frame"))[, -1]
trump_dtm2_dt <- colSums(trump_dtm2_dt)
trump_dtm2_dt <- data.table(value = unname(trump_dtm2_dt),
                            term = names(trump_dtm2_dt))
trump_dtm2_dt[, c("term1", "term2") := tstrsplit(term, "_", fixed = TRUE)]
trump_dtm2_dt$term <- NULL
trump_dtm2_dt <- trump_dtm2_dt[, c("term1", "term2", "value")]

bigram_graph <- igraph::graph_from_data_frame(trump_dtm2_dt[value > 200, ])
set.seed(2017)
ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1)
```


### How similar are speeches before and after Trump became the president? 
Cluster analysis of similarty between speeches in 2018 and 2019.
```{r}
## ------------------------------------------
## hierarchical clustering - get distances on normalized dfm
tstat_dist <- textstat_dist(dfm_weight(trump_dfm, scheme = "prop")) %>%
  as.dist()

pres_cluster <- hclust(tstat_dist)
# label with document names
pres_cluster$labels <- trump$year

dend <- as.dendrogram(pres_cluster)
colors_to_use <- as.numeric(trump$year)
colors_to_use <- colors_to_use[order.dendrogram(dend)]
labels_colors(dend) <- colors_to_use
# plot as a dendrogram
plot(dend, xlab = "", sub = "", 
     main = "Similarity of Speeches by Year", 
     cex = .6)
legend("topright", legend = c("2018", "2019"), col = c("red", "black"), pch = 19)
```


### Has the frequency of talking about migration changed in time? 
Time series analysis of some key campaign topics.
```{r}
## time series immigration
trump_dtm <- as.data.table(convert(trump_dfm, to = "data.frame"))[, -1]
trump_dtm$date <- trump$upload_date
trump_dtm[, year_month := format(date, format = "%Y-%m")]
immigrants_summary <- trump_dtm[, .(immigration_count = sum(immigration)), by = year_month]
immigrants_summary <- immigrants_summary[order(as.numeric(gsub("-","",year_month))), ]

plot(immigrants_summary$immigration_count, type = "l", xaxt = "n",
     ylab = "Count", main = "Count of word Immigration", xlab = "")
points(immigrants_summary$immigration_count, pch = 16)
axis(1, at=1:nrow(immigrants_summary), 
     labels=immigrants_summary$year_month, 
     las = 2, cex.axis = 0.8)
```

```{r}
## time series health care
health_care_summary <- trump_dtm[, .(hc_count = sum(health + care)), by = year_month]
health_care_summary <- health_care_summary[order(as.numeric(gsub("-","",year_month))), ]

plot(health_care_summary$hc_count, type = "l", xaxt = "n",
     ylab = "Count", main = "Count of word Health and Care", xlab = "")
points(health_care_summary$hc_count, pch = 16)
axis(1, at=1:nrow(health_care_summary), 
     labels=health_care_summary$year_month, 
     las = 2, cex.axis = 0.8)
```


```{r}
## time series trade
trade_summary <- trump_dtm[, .(count = sum(trade)), by = year_month]
trade_summary <- trade_summary[order(as.numeric(gsub("-","",year_month))), ]

plot(trade_summary$count, type = "l", xaxt = "n",
     ylab = "Count", main = "Count of word Trade", xlab = "")
points(trade_summary$count, pch = 16)
axis(1, at=1:nrow(trade_summary), 
     labels=trade_summary$year_month, 
     las = 2, cex.axis = 0.8)
```


```{r}
## time series trade
jobs_summary <- trump_dtm[, .(count = sum(job + jobs)), by = year_month]
jobs_summary <- jobs_summary[order(as.numeric(gsub("-","",year_month))), ]

plot(jobs_summary$count, type = "l", xaxt = "n",
     ylab = "Count", main = "Count of word Jobs", xlab = "")
points(jobs_summary$count, pch = 16)
axis(1, at=1:nrow(jobs_summary), 
     labels=jobs_summary$year_month, 
     las = 2, cex.axis = 0.8)
```